Computing properties of independent sets by generic programming tensor networks \thanks{\funding{...}}

### 2021-09-11 paper/paper.tex
complete benchmark

\begin{example}
$C_{ijk} = A_{jkm} B_{mil} V_{jm}$ is a tensor network representing $C_{ijk} = \sum_{ml}A_{jkm} B_{mil} V_{jm}$ 
\red{I've changed this. please check. The original was $C_{ijk} = A_{jkm} B_{mil} V_{jm}$ and $C_{ijk} = \sum_{ml}A_{jkm} B_{mil} V_{jm}$}.
\blue{changed back because $l$ is not an open edge. Is it still confusing if I introduce open circles for open edges?}
Its hypergraph representation is shown below, where we use different colors to annotate different hyperedges and open circles to denote open edges.

\centerline{\begin{tikzpicture}[
    dot/.style = {circle, fill, minimum size=#1,
                inner sep=0pt, outer sep=0pt},
    dot/.default = 6pt  % size of the circle diameter 
                    ]  
    \def\dx{0};
    \def\r{0.5cm}
    \def\sr{0.15cm}
    \def\ax{0}
    \def\ay{0}
    \def\bx{1}
    \def\by{1}
    \def\cx{1}
    \def\cy{-1}
    \node[color=white,fill=black,dot=\r] at (\ax+\dx,\ax) (a) {A};
    \node[color=white,fill=black,dot=\r] at (\bx+\dx,\by) (b) {B};
    \node[color=white,fill=black,dot=\r] at (\cx+\dx,\cy) (v) {V};
    \node[color=white,draw=black,dot=\sr] at (\ax-1,\by) (o1) {};
    \node[color=white,draw=black,dot=\sr] at (\ax-1.5,\ay) (o2) {};
    \node[color=white,draw=black,dot=\sr] at (\ax-1,\cy) (o3) {};
    \node at (\ax-0.8,\ay) (k) {k};
    \node at (\bx+0.4,\ay) (m) {m};
    \node at (\ax,\cy) (j) {j};
    \node at (\bx+1,\by) (l) {l};
    \node at (\bx-1,\by) (i) {i};
    \draw[color=blue,thick] (i) -- (b);
    \draw[color=blue,thick] (i) -- (o1);
    \draw[color=cyan,thick] (l) -- (b);
    \draw[color=violet,thick] (k) -- (a);
    \draw[color=violet,thick] (k) -- (o2);
    \draw[color=black,thick] (b) -- (m);
    \draw[color=black,thick] (m) -- (a);
    \draw[color=black,thick] (m) -- (v);
    \draw[color=red,thick] (a) -- (j);
    \draw[color=red,thick] (v) -- (j);
    \draw[color=red,thick] (o3) -- (j);
\end{tikzpicture}}
\end{example}

In the main text, we use the generalized tensor network notation. The generalized tensor network can be easily translated to the standard notation by adding identity tensors denoted as $\delta$ tensors.
The example above can be written as $C_{ijkl} = A_{jkm} B_{mil} V_{jm} = A_{ukm} B_{pil} V_{vq} \delta_{mpq} \delta_{juv}$ \red{update this} in the standard notion. However, by adding the $\delta$ tensors, one may unnecessarily increase the contraction complexity of a graph, which we illustrate in \App{app:tensorbad}.

\section{Generic programming}\label{sec:generic}
\red{I suggest we have a section on generic programming, putting both the semiring stuff and the summary table to this section, since generic programming is one of the highlights of this paper as well. I think it doesn't need to be long. Jinguo, could you write one or two paragraphs introducing generic programming?}\blue{I wrote a draft here, does it look ok in general?}
In previous works combining tensor network and combinatoric problems~\cite{Kourtis2019,Biamonte2017},
the elements in tensor networks are limited to regular number types such as floating point numbers and integers.
As the development of modern compiling technology, we do not need to limit our imaginations to regular number types anymore.
The key technology playing the role is called the generic programming:

@@

The independence polynomial is closely related to the matching polynomial, the clique polynomial, and the vertex cover polynomial.
In fact, the independence polynomial can be viewed as a generalization of the matching polynomial, since the matching polynomial of a graph $G$ is the same as the independence polynomial of the line graph of $G$~\cite{Levit2005}.
Thus, our algorithm to compute the independence polynomial can also be used to compute the above graph polynomials. Some properties of the independence polynomial, such as the uni-modality, log-concavity, and roots of the polynomial, are well studied in the literature for special classes of graphs~\cite{Levit2005}.
We hope our algorithms to calculate the independence polynomial can also help further research in these directions. \red{possibly add set packing problem relationship}

To compute the independence polynomial of a graph $G$, we encode it to a tensor network. On the vertex $i$ of the graph $G$, we place a rank-one vertex tensor, $W(x_i)$, of size $2$ parametrized by $x_{i}$ and a rank-two edge tensor, $B$, of size $2 \times 2$ on an edge $(i,j)$
\red{I removed the indices $s_{i}$ etc, on the right of the equations, since that looks confusing to me. If I understand correctly, $s_{i}, s_{j}$ just denotes the index of the vector/matrix.}
\blue{To be consistent, we should also remove the index on the LHS. It is better not removing the label information.}
\begin{equation}
    W(x_i)_{s_i} = \left(\begin{matrix}
        1 \\
        x_i
    \end{matrix}\right),
    \qquad \quad 
       B_{s_i s_j} = \left(\begin{matrix}
        1  & 1\\
        1 & 0
    \end{matrix}\right), \label{eq:tensor}
\end{equation} \red{remove subscript $s_{i} s_{j}$ and add description after the next equation}
where the tensor index $s_i$ is a boolean variable; $s_{i} = 1$ corresponds that the vertex $i$ is in the independent set, and $s_{i} = 0$ means otherwise. Let us assume the tensor elements are real numbers for now.
The edge tensor element $B_{s_{i}=1, s_{j}=1} = 0$ encodes the independent set constraint, meaning vertex $i$ and $j$ cannot be both in the independent set if they are connected by an edge $(i,j)$. The contraction of the tensor network representing $G$ gives
\begin{equation}
    P(G, \{x_1, x_{2}, \ldots,x_{|V|}\}) = \sum\limits_{s_1, s_2, \ldots, s_{|V|} = 0}^{1} \prod\limits_{i=1}^{|V|} W(x_i)_{s_i} \prod\limits_{(i,j) \in E(G)} B_{s_i s_j},
\end{equation}
where the summation runs over all vertex configurations $\{s_1, s_{2}, \ldots,s_{|V|}\}$ and accumulates the product of tensor elements to the scalar output $P$ (see Example \ref{eg:tensorcontraction} for a concrete example).
In the special case of $x_i = x$, the contraction result directly corresponds to the independence polynomial.
The connection can be understood as follows: the product over vertex tensor elements produces a factor $x^k$, where $k=\sum_i s_i$ counts the set size,
and the product over edge tensor elements gives a factor $1$ for a configuration being in an independent set and $0$ otherwise. In the implementation of the algorithm, we benefit from recent advances in tensor network contraction,
where researchers working on quantum circuit simulation evaluate the tensor network by pairwise contracting tensors in a good heuristic order~\cite{Gray2021, Pan2021}.
A good contraction order can reduce the time complexity significantly, at the cost of having a space overhead of $O(2^{\text{tw}(G)})$~\cite{Markov2008}.
The pairwise tensor contraction also makes it possible to utilize basic linear algebra subprograms (BLAS) functions to speed up the computation for certain tensor element types.

\begin{example}\label{eg:tensorcontraction}
Mapping a graph (left) to a tensor network (right) that encodes the independence polynomial. In the generalized tensor network's graphical representation, a vertex is mapped to a hyperedge, and a vertex tensor is connected to the hyperedge. An edge is mapped to an edge tensor connected to the hyperedge.
\red{come back to this to change}\blue{$\checkmark$}

@@

Here, we propose to find the independence polynomial by fitting $\alpha(G)+1$ random values of $x_{i}$ and $y_{i} = I(G,x_{i})$. One can then compute the independence polynomial coefficients $a_{i}$ by solving the linear equation: 
\begin{equation}
\left(\begin{matrix}
1 & x_0 & x_0^2 & \ldots & x_0^{\alpha(G)} \\
1 & x_1 & x_1^2 & \ldots & x_1^{\alpha(G)} \\
\vdots & \vdots & \vdots &\ddots & \vdots \\
1 & x_{\alpha(G)} & x_{\alpha(G)}^2 & \ldots & x_{\alpha(G)}^{\alpha(G)}
\end{matrix}\right)
\left(\begin{matrix}
a_0 \\ a_1 \\ \vdots \\ a_{\alpha(G)}
\end{matrix}\right)
= \left(\begin{matrix}
y_0 \\ y_1 \\ \vdots \\ y_{\alpha(G)}
\end{matrix}\right).\label{eq:lineareq}
\end{equation}
With this approach, we do not incur the linear overhead in space. However, because the independence polynomial coefficients can have a huge order-of-magnitude range, if we use floating point numbers in the computation, the round-off errors can be significant for the counting of large independent set sizes.
In addition, the number could easily overflow if we use fixed-width integer types.
The big integer type is also not a good option because big integers with varying width can be very slow and is incompatible with graphics processing unit (GPU) devices. These problems can be solved by introducing a finite-field algebra $\text{GF}(p)$:
\begin{equation}
\begin{split}
    x ~\oplus~ y &= x+y\pmod p,\\
    x ~\odot~ y &= xy\pmod p,\\
    \mymathbb{0} &= 0,\\
    \mymathbb{1} &= 1.
\end{split}\label{eq:finitefield}
\end{equation}
With a finite-field algebra, we have the following observations:
\begin{enumerate}
    \item One can use Gaussian elimination~\cite{Golub2013} to solve the linear equation \Eq{eq:lineareq} since it is a generic algorithm that works for any elements with field algebra. The multiplicative inverse of a finite-field algebra can be computed with the extended Euclidean algorithm.
    \item Given the remainders of a larger unknown integer $x$ over a set of co-prime integers $\{p_1, p_2, \ldots, p_n\}$,
    $x \pmod {p_1 \times p_2 \times \ldots \times p_n}$ can be computed using the Chinese remainder theorem. With this, one can infer big integers from small integers.
\end{enumerate}
With these observations, we develop Algorithm~\ref{alg:finitefield} to compute the independence polynomial exactly without introducing space overheads.
This is an iterative algorithm that tries different prime numbers $p$ until convergence.
In each iteration, we contract tensor networks to evaluate the polynomial for each $x\in X$ on the finite field algebra ${\rm GF}(p)$.
Let use denote the results as $Y$, then we solve a linear equation $Y = X A_p$ using the gaussian elimination on ${\rm GF}(p)$ to find the coefficient moduluses $A_p \equiv A \pmod p$.
Finally we apply the Chinese remainder theorem to get $A \pmod P$, where $P$ is a product of all prime numbers chosen in current and previous iterations.
If this number does not change comparing with the previous iteration, it indicates a convergent result.
All computations are done with integers of fixed width $W$ except the last step of applying Chinese remainder theorem.
In \App{app:fft}, we provide another method to solve the linear equation using discrete Fourier transform.

\begin{algorithm}[!ht]
    \small
    \SetAlgoNoLine
    \LinesNumbered
    Let $P = 1$, $W$ be the integer width, vector $\chi = (0,1,2, \ldots, \alpha(G))$, matrix $X_{ij} = (\chi_i)^j$, where $i,j = 0, 1, \ldots, \alpha(G)$\;

@@

        compute the tensor contraction for each $x = \chi_i \pmod p$ on $\text{GF}(p)$ and obtain $Y = (y_0, y_1, \ldots , y_{\alpha(G)}) \pmod p $\; \red{for each $i$}

@@

The variable $x_{i}$ in the vertex tensor is initialized to $x_i = \{\boldsymbol{e}_{i}\}$, where $\boldsymbol{e}_i$ is the standard basis vector of size $|V|$ and having 1 at index $i$. The vertex and edge tensors are thus
\begin{equation}
    W^{\rm enum}(\{\boldsymbol{e}_{i}\}) = \left(\begin{matrix}
        \mymathbb{1} \\
        \{\boldsymbol{e}_{i}\}
    \end{matrix}\right),   
    \qquad \qquad
        B^{\rm enum} = \left(\begin{matrix}
        \mymathbb{1}  & \mymathbb{1} \\
        \mymathbb{1} & \mymathbb{0}
    \end{matrix}\right).
\end{equation}
Contraction of the tensor network enumerates all the independent sets.
To enumerate only the MISs, we can combine the above algebra \Eq{eq:set} with the tropical algebra in \Eq{eq:countingtropical} and define $s(k) = (s_{k}, k)$, where the first element follows the algebra in \Eq{eq:set} and the second element follows the max-plus tropical algebra.
The combined operations become: 
\begin{equation}
\begin{split}
    s(k) \oplus s(j) &= \begin{cases}
        \left(s_k \cup s_j, \max(k,j) \right), & k = j \\
        \left(s_j, \max(k,j) \right), & k < j \\
        \left(s_k, \max(k,j) \right), & k > j
    \end{cases}, \\
    s(k) \odot s(j) &= (\{\sigma \lor^\circ \tau \, | \, \sigma \in s_k, \tau \in s_j\}, k+j) \\
    \mymathbb{0} &= (\{ \}, -\infty)\\
    \mymathbb{1} &= (\{0^{\otimes |V|}\}, 0). \label{eq:countingtropicalset}
\end{split}
\end{equation}
%\begin{equation}
%\begin{split}
%    s_x\infty^x \oplus s_y\infty^y &= \begin{cases}
%        (s_x \cup s_y)\infty^{\max(x,y)}, & x = y\\
%        s_y\infty^{\max(x,y)}, & x < y\\
%        s_x\infty^{\max(x,y)}, & x > y
%    \end{cases},\\
%    s_x\infty^x \odot s_y\infty^y &= \{\sigma \lor^\circ \tau | \sigma \in s_x, \tau \in s_y\}\infty^{x+y},\\
%    \mymathbb{0} &= \{\}\infty^{-\infty},\\
%    \mymathbb{1} &= \{0^{\otimes |V|}\}\infty^{0},
%\end{split}
%\end{equation}
Clearly, the vertex tensor and edge tensor become
\begin{equation}
    W^{\rm MISenum}(\{\boldsymbol{e}_{i}\}) = \left(\begin{matrix}
        \mymathbb{1} \\
        \left({\{\boldsymbol{e}_{i}\}, 1} \right)
    \end{matrix}\right),   
    \qquad \qquad
        B^{\rm MISenum} = \left(\begin{matrix}
        \mymathbb{1}  & \mymathbb{1} \\
        \mymathbb{1} & \mymathbb{0}
    \end{matrix}\right).
\end{equation}
The contraction of the corresponding tensor network yields an enumeration of all MIS configurations.
If one is interested in obtaining only a single MIS configuration, one can just keep a single configuration in the intermediate computations to save the computational effort.
Here is a new algebra defined on the bit strings, replacing the sets of bit strings in \Eq{eq:set}, 
%We leave this as an exercise for readers.
%
%\iffalse
%\begin{equation}
%\begin{split}
%    \sigma_x\infty^x \oplus \sigma_y\infty^y &= \begin{cases}
%        {\rm select}(\sigma_x, \sigma_y)\infty^{\max(x,y)}, & x = y\\
%        \sigma_y\infty^{\max(x,y)}, & x < y\\
%        \sigma_x\infty^{\max(x,y)}, & x > y
%    \end{cases},\\
%    \sigma_x\infty^x \odot \sigma_y\infty^y &= (\sigma_x \lor^\circ \sigma_y)\infty^{x+y},\\
%    \mymathbb{0} &= 1^{\otimes n}\infty^{-\infty},\\
%    \mymathbb{1} &= 0^{\otimes n}\infty^{0},
%\end{split}
%\end{equation}
%\fi
\begin{equation}
\begin{split}
    \sigma \oplus \tau &= {\rm select}(\sigma, \tau), \\
    \sigma \odot \tau &= (\sigma\lor^\circ \tau),\\
    \mymathbb{0} &= 1^{\otimes |V|}, \\
    \mymathbb{1} &= 0^{\otimes |V|},
\end{split}\label{eq:singleconfig}
\end{equation}
where the \texttt{select} function picks one of $\sigma$ and $\tau$ by some criteria to make the algebra commutative and associative, e.g. by picking one with a smaller integer value.
%In practice, we can just pick randomly from them, in which case the program will output one of the MIS configurations randomly.
\red{Can we also enumerate all maximal independent sets by combining with \Eq{eq:maximal} or enumerate suboptimal solutions by combining with the truncated polynomial?}

@@

\begin{figure} 
    \centering
    \includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm}, clip]{figures/fig1.pdf}
    \caption{Benchmark results for computing different properties of independent sets on a random three regular graph with different tensor element types.
    (b) is the computing time for computing maximum independent set size, number of independent sets and suboptimal sets.
    (c) is the computing time for independence polynomials with different approaches.
    (d) is the computing time for finding all configurations and a single configuration, with or without bounding.
    }\red{split into two graphs: one with time versus number of vertices, $|V|$, and one with time versus tree width.}
    \red{We also need to show/say what types of graphs is used here for benchmarking.}
    \label{fig:benchmark}
\end{figure}

Among the benchmarks shown in \Fig{fig:benchmark}, computing \texttt{max size (CPU)},
\texttt{total size (CPU)}, \texttt{total size (GPU)}, \texttt{IDP (FFT) (CPU)}, \texttt{IDP (FFT) (GPU)} and \texttt{single configuration (bounding) (CPU)} can utilize fast BLAS libraries,
hence they are relatively faster comparing to other methods in the same category.
Only algebras requiring dynamic sized structures can not be computed on GPU, here are polynomial algebra for computing independence polynomial and set algebra for enumerating configurations.
When computing the independence polynomial, the finite field approach is the only method without round off error that can run on GPU.

@@

@@

>>>>

\begin{example}
$C_{ijk} = A_{jkm} B_{mil} V_{jm}$ is a tensor network representing $C_{ijk} = \sum_{ml}A_{jkm} B_{mil} V_{jm}$,
where $i,j,k$ in the output tensor are open edges.
Its hypergraph representation is shown below, where we use different colors to annotate different hyperedges and open circles to denote open edges.

\centerline{\begin{tikzpicture}[
    dot/.style = {circle, fill, minimum size=#1,
                inner sep=0pt, outer sep=0pt},
    dot/.default = 6pt  % size of the circle diameter 
                    ]  
    \def\dx{0};
    \def\r{0.5cm}
    \def\sr{0.15cm}
    \def\ax{0}
    \def\ay{0}
    \def\bx{1}
    \def\by{1}
    \def\cx{1}
    \def\cy{-1}
    \node[color=white,fill=black,dot=\r] at (\ax+\dx,\ax) (a) {A};
    \node[color=white,fill=black,dot=\r] at (\bx+\dx,\by) (b) {B};
    \node[color=white,fill=black,dot=\r] at (\cx+\dx,\cy) (v) {V};
    \node[color=transparent,draw=transparent,dot=0] at (\ax-1,\by) (o1) {};
    \node[color=transparent,draw=transparent,dot=0] at (\ax-1.5,\ay) (o2) {};
    \node[color=transparent,draw=transparent,dot=0] at (\ax-1,\cy) (o3) {};
    \node at (\ax-0.8,\ay) (k) {k};
    \node at (\bx+0.4,\ay) (m) {m};
    \node at (\ax,\cy) (j) {j};
    \node at (\bx+1,\by) (l) {l};
    \node at (\bx-1,\by) (i) {i};
    \draw[color=blue,thick] (i) -- (b);
    \draw[color=blue,thick] (i) -- (o1);
    \draw[color=cyan,thick] (l) -- (b);
    \draw[color=violet,thick] (k) -- (a);
    \draw[color=violet,thick] (k) -- (o2);
    \draw[color=black,thick] (b) -- (m);
    \draw[color=black,thick] (m) -- (a);
    \draw[color=black,thick] (m) -- (v);
    \draw[color=red,thick] (a) -- (j);
    \draw[color=red,thick] (v) -- (j);
    \draw[color=red,thick] (o3) -- (j);
\end{tikzpicture}}
\end{example}

In the main text, we use the generalized tensor network notation. The generalized tensor network can be easily translated to the standard notation by adding identity tensors denoted as $\delta$ tensors.
The example above can be written as $C_{ijk} = A_{jkm} B_{mil} V_{jm} = A_{ukm} B_{pil} V_{vq} \delta_{mpq} \delta_{juv}\delta_l$ in the standard notion. However, by adding the $\delta$ tensors, one may unnecessarily increase the contraction complexity of a graph, which we illustrate in \App{app:tensorbad}.

\section{Generic programming}\label{sec:generic}
In previous works combining tensor network and combinatoric problems~\cite{Kourtis2019,Biamonte2017},
the elements in tensor networks are limited to regular number types such as floating point numbers and integers.
As the development of modern compiling technology, we do not need to limit our imaginations to regular number types anymore.
The key technology playing the role is called the generic programming:

@@

The independence polynomial is closely related to the matching polynomial, the clique polynomial, and the vertex cover polynomial.
In fact, the independence polynomial can be viewed as a generalization of the matching polynomial, since the matching polynomial of a graph $G$ is the same as the independence polynomial of the line graph of $G$~\cite{Levit2005}.
Thus, our algorithm to compute the independence polynomial can also be used to compute the above graph polynomials. Some properties of the independence polynomial, such as the uni-modality, log-concavity, and roots of the polynomial, are well studied in the literature for special classes of graphs~\cite{Levit2005}.
We hope our algorithms to calculate the independence polynomial can also help further research in these directions. \red{possibly add set packing problem relationship} \blue{I added a subsection in \App{app:otherproblems}?}

To compute the independence polynomial of a graph $G$, we encode it to a tensor network. On the vertex $i$ of the graph $G$, we place a rank-one vertex tensor, $W(x_i)$, of size $2$ parametrized by $x_{i}$ and a rank-two edge tensor, $B$, of size $2 \times 2$ on an edge
\red{I removed the indices $s_{i}$ etc, on the right of the equations, since that looks confusing to me. If I understand correctly, $s_{i}, s_{j}$ just denotes the index of the vector/matrix.}
\blue{To be consistent, we should also remove the index on the LHS. It is better not removing the label information.}
\blue{now?}
\begin{equation}
    W(x_i) = \left(\begin{matrix}
        1 \\
        x_i
    \end{matrix}\right),
    \qquad \quad 
       B = \left(\begin{matrix}
        1  & 1\\
        1 & 0
    \end{matrix}\right), \label{eq:tensor}
\end{equation}
The corresponding tensor network contraction gives
\begin{equation}
    P(G, \{x_1, x_{2}, \ldots,x_{|V|}\}) = \sum\limits_{s_1, s_2, \ldots, s_{|V|} = 0}^{1} \prod\limits_{i=1}^{|V|} W(x_i)_{s_i} \prod\limits_{(i,j) \in E(G)} B_{s_i s_j},
\end{equation}
where the summation runs over all vertex configurations $\{s_1, s_{2}, \ldots,s_{|V|}\}$ and accumulates the product of tensor elements to the output $P$ (see Example \ref{eg:tensorcontraction} for a concrete example).
The tensor index $s_i$ is a boolean variable; $s_{i} = 1$ corresponds that the vertex $i$ is in the independent set, and $s_{i} = 0$ means otherwise. Let us assume the tensor elements are real numbers for now.
The edge tensor element $B_{s_{i}=1, s_{j}=1} = 0$ encodes the independent set constraint, meaning vertex $i$ and $j$ cannot be both in the independent set if they are connected by an edge $(i,j)$. The contraction of the tensor network representing $G$ gives
In the special case of $x_i = x$, the contraction result directly corresponds to the independence polynomial.
The connection can be understood as follows: the product over vertex tensor elements produces a factor $x^k$, where $k=\sum_i s_i$ counts the set size,
and the product over edge tensor elements gives a factor $1$ for a configuration being in an independent set and $0$ otherwise. In the implementation of the algorithm, we benefit from recent advances in tensor network contraction,
where researchers working on quantum circuit simulation evaluate the tensor network by pairwise contracting tensors in a good heuristic order~\cite{Gray2021, Pan2021}.
A good contraction order can reduce the time complexity significantly, at the cost of having a space overhead of $O(2^{\text{tw}(G)})$~\cite{Markov2008}.
The pairwise tensor contraction also makes it possible to utilize basic linear algebra subprograms (BLAS) functions to speed up the computation for certain tensor element types.

\begin{example}\label{eg:tensorcontraction}
Mapping a graph (left) to a tensor network (right) that encodes the independence polynomial.
In the generalized tensor network's graphical representation, a vertex is mapped to a hyperedge.
We attach a vertex tensor on each hyperedge and an edge tensor between two hyperedges if vertices of the two hyperedges are connected in the original graph.
\red{come back to this to change}\blue{now?}

@@

Here, we propose to find the independence polynomial by fitting $\alpha(G)+1$ random values of $x_{i}$ and $y_{i} = I(G,x_{i})$. One can then compute the independence polynomial coefficients $a_{i}$ by solving the linear equation: 
\begin{equation}
\left(\begin{matrix}
1 & x_0 & x_0^2 & \ldots & x_0^{\alpha(G)} \\
1 & x_1 & x_1^2 & \ldots & x_1^{\alpha(G)} \\
\vdots & \vdots & \vdots &\ddots & \vdots \\
1 & x_{\alpha(G)} & x_{\alpha(G)}^2 & \ldots & x_{\alpha(G)}^{\alpha(G)}
\end{matrix}\right)
\left(\begin{matrix}
a_0 \\ a_1 \\ \vdots \\ a_{\alpha(G)}
\end{matrix}\right)
= \left(\begin{matrix}
y_0 \\ y_1 \\ \vdots \\ y_{\alpha(G)}
\end{matrix}\right).\label{eq:lineareq}
\end{equation}
With this approach, we do not incur the linear overhead in space. However, because the independence polynomial coefficients can have a huge order-of-magnitude range, if we use floating point numbers in the computation, the round-off errors can be significant for the counting of large independent set sizes.
In addition, the number could easily overflow if we use fixed-width integer types.
The big integer type is also not a good option because big integers with varying width can be very slow and is incompatible with graphics processing unit (GPU) devices. These problems can be solved by introducing a finite-field algebra $\text{GF}(p)$:
\begin{equation}
\begin{split}
    x ~\oplus~ y &= x+y\pmod p,\\
    x ~\odot~ y &= xy\pmod p,\\
    \mymathbb{0} &= 0,\\
    \mymathbb{1} &= 1.
\end{split}\label{eq:finitefield}
\end{equation}
With a finite-field algebra, we have the following observations:
\begin{enumerate}
    \item One can use Gaussian elimination~\cite{Golub2013} to solve the linear equation \Eq{eq:lineareq} since it is a generic algorithm that works for any elements with field algebra. The multiplicative inverse of a finite-field algebra can be computed with the extended Euclidean algorithm.
    \item Given the remainders of a larger unknown integer $x$ over a set of co-prime integers $\{p_1, p_2, \ldots, p_n\}$,
    $x \pmod {p_1 \times p_2 \times \ldots \times p_n}$ can be computed using the Chinese remainder theorem. With this, one can infer big integers from small integers.
\end{enumerate}
With these observations, we develop Algorithm~\ref{alg:finitefield} to compute the independence polynomial exactly without introducing space overheads.
This is an iterative algorithm that tries different prime numbers $p$ until convergence.
In each iteration, we contract tensor networks to evaluate the polynomial for each $x\in X$ on the finite field algebra ${\rm GF}(p)$.
Let use denote the results as $Y$, then we solve a linear equation $Y = X A_p$ using the gaussian elimination on ${\rm GF}(p)$ to find the coefficient moduluses $A_p \equiv A \pmod p$.
Finally we apply the Chinese remainder theorem to get $A \pmod {P\times p}$, where $P$ is a product of all prime numbers chosen in previous iterations.
If this number does not change comparing with the previous iteration, it indicates a convergent result.
All computations are done with integers of fixed width $W$ except the last step of applying Chinese remainder theorem.
In \App{app:fft}, we provide another method to solve the linear equation using discrete Fourier transform.

\LinesNumberedHidden
\begin{algorithm}[!ht]
    \small
    \SetAlgoNoLine
    %\LinesNumbered
    Let $P = 1$, $W$ be the integer width, vector $\chi = (0,1,2, \ldots, \alpha(G))$, matrix $X_{ij} = (\chi_i)^j$, where $i,j = 0, 1, \ldots, \alpha(G)$\;

@@

        \For{$i=1\ldots\alpha(G)$}{
            t = ${\rm tensor\_network}(\chi_i\pmod p)$ \tcp*[l]{defined on $\text{GF}(p)$}

            $Y_i \pmod p  = {\rm contract}(t)$\;
        }

@@

The variable $x_{i}$ in the vertex tensor is initialized to $x_i = \{\boldsymbol{e}_{i}\}$, where $\boldsymbol{e}_i$ is the standard basis vector of size $|V|$ and having 1 at index $i$. The vertex and edge tensors are thus
\begin{equation}
    W^{\rm enum}(\{\boldsymbol{e}_{i}\}) = \left(\begin{matrix}
        \mymathbb{1} \\
        \{\boldsymbol{e}_{i}\}
    \end{matrix}\right),   
    \qquad \qquad
        B^{\rm enum} = \left(\begin{matrix}
        \mymathbb{1}  & \mymathbb{1} \\
        \mymathbb{1} & \mymathbb{0}
    \end{matrix}\right).
\end{equation}

This sets algebra can serve as the coefficients in \Eq{eq:countingtropical} to enumerate all MISs, \Eq{eq:polynomial} to enumerate independent sets of different sizes,
or \Eq{eq:max2poly} to enumerate all independent sets of size $\alpha(G)$ and $\alpha(G)-1$.
For example, to enumerate only the MISs, with the tropical algebra, we define $s(k) = (s_{k}, k)$, where the first element follows the algebra in \Eq{eq:set} and the second element follows the max-plus tropical algebra.
The combined operations become: 
\begin{equation}
\begin{split}
    s(k) \oplus s(j) &= \begin{cases}
        \left(s_k \cup s_j, \max(k,j) \right), & k = j \\
        \left(s_j, \max(k,j) \right), & k < j \\
        \left(s_k, \max(k,j) \right), & k > j
    \end{cases}, \\
    s(k) \odot s(j) &= (\{\sigma \lor^\circ \tau \, | \, \sigma \in s_k, \tau \in s_j\}, k+j) \\
    \mymathbb{0} &= (\{ \}, -\infty)\\
    \mymathbb{1} &= (\{0^{\otimes |V|}\}, 0). \label{eq:countingtropicalset}
\end{split}
\end{equation}
%\begin{equation}
%\begin{split}
%    s_x\infty^x \oplus s_y\infty^y &= \begin{cases}
%        (s_x \cup s_y)\infty^{\max(x,y)}, & x = y\\
%        s_y\infty^{\max(x,y)}, & x < y\\
%        s_x\infty^{\max(x,y)}, & x > y
%    \end{cases},\\
%    s_x\infty^x \odot s_y\infty^y &= \{\sigma \lor^\circ \tau | \sigma \in s_x, \tau \in s_y\}\infty^{x+y},\\
%    \mymathbb{0} &= \{\}\infty^{-\infty},\\
%    \mymathbb{1} &= \{0^{\otimes |V|}\}\infty^{0},
%\end{split}
%\end{equation}
Clearly, the vertex tensor and edge tensor become
\begin{equation}
    W^{\rm MISenum}(\{\boldsymbol{e}_{i}\}) = \left(\begin{matrix}
        \mymathbb{1} \\
        \left({\{\boldsymbol{e}_{i}\}, 1} \right)
    \end{matrix}\right),   
    \qquad \qquad
        B^{\rm MISenum} = \left(\begin{matrix}
        \mymathbb{1}  & \mymathbb{1} \\
        \mymathbb{1} & \mymathbb{0}
    \end{matrix}\right).
\end{equation}
The contraction of the corresponding tensor network yields an enumeration of all MIS configurations.

If one is interested in obtaining only a single MIS configuration, one can just keep a single configuration in the intermediate computations to save the computational effort.
Here is a new algebra defined on the bit strings, replacing the sets of bit strings in \Eq{eq:set}, 
%We leave this as an exercise for readers.
%
%\iffalse
%\begin{equation}
%\begin{split}
%    \sigma_x\infty^x \oplus \sigma_y\infty^y &= \begin{cases}
%        {\rm select}(\sigma_x, \sigma_y)\infty^{\max(x,y)}, & x = y\\
%        \sigma_y\infty^{\max(x,y)}, & x < y\\
%        \sigma_x\infty^{\max(x,y)}, & x > y
%    \end{cases},\\
%    \sigma_x\infty^x \odot \sigma_y\infty^y &= (\sigma_x \lor^\circ \sigma_y)\infty^{x+y},\\
%    \mymathbb{0} &= 1^{\otimes n}\infty^{-\infty},\\
%    \mymathbb{1} &= 0^{\otimes n}\infty^{0},
%\end{split}
%\end{equation}
%\fi
\begin{equation}
\begin{split}
    \sigma \oplus \tau &= {\rm select}(\sigma, \tau), \\
    \sigma \odot \tau &= (\sigma\lor^\circ \tau),\\
    \mymathbb{0} &= 1^{\otimes |V|}, \\
    \mymathbb{1} &= 0^{\otimes |V|},
\end{split}\label{eq:singleconfig}
\end{equation}
where the \texttt{select} function picks one of $\sigma$ and $\tau$ by some criteria to make the algebra commutative and associative, e.g. by picking one with a smaller integer value.
%In practice, we can just pick randomly from them, in which case the program will output one of the MIS configurations randomly.
\red{Can we also enumerate all maximal independent sets by combining with \Eq{eq:maximal} or enumerate suboptimal solutions by combining with the truncated polynomial?}
\blue{now?}

Although the above discussion is in the context of the the MIS problem, we can also combine the above algebras with the maximal independence network in \Eq{eq:maximal} to enumerate or sample maximal independent sets,
as well as other problems in \App{app:otherproblems} for configuration enumeration and sampling.

@@

\begin{figure} 
    \centering
    \includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm}, clip]{figures/fig1.pdf}
    \caption{Benchmark results for computing different properties of independent sets on a random three regular graph with different tensor element types.
    (b) is the computing time for computing maximum independent set size, number of independent sets and suboptimal sets.
    (c) is the computing time for independence polynomials with different approaches.
    (d) is the computing time for finding all configurations and a single configuration, with or without bounding.
    }
    \label{fig:benchmark}
\end{figure}

Among the benchmarks shown in \Fig{fig:benchmark}, computing \texttt{max size (CPU)},
\texttt{total size (CPU)}, \texttt{total size (GPU)}, \texttt{IDP (FFT) (CPU)}, \texttt{IDP (FFT) (GPU)} and \texttt{single configuration (bounding) (CPU)} can utilize fast BLAS functions,
hence they are relatively faster comparing to other methods in the same category.
Only algebras requiring dynamic sized structures can not be computed on GPU, here are polynomial algebra for computing independence polynomial and set algebra for enumerating configurations.
When computing the independence polynomial, the finite field approach is the only method without round off error that can run on GPU.
To enumerate the MISs, the technique to bound the enumeration space is important,
without which the largest enumeratible size is only $150$ with 32G memory while being one order slower.

@@

\subsection{Max cut problem}
Max cut problem is also known as the boolean spinglass problem.
Its tensor network representation does not contain vertex tensors.
The edge tensor can be defined as
\begin{equation}
    B(x_{ij}) = \left(\begin{matrix}
        1 & x_{ij}\\
        x_{ij} & 1
    \end{matrix}\right).
\end{equation}
We can also define a polynomial about edges variables by setting $x_{ij} = x$,
its $k$th coefficient is two times the number of configurations of cut size $k$.

\subsection{Set packing}
Set packing the the hypergraph version of the maximum independent set problem, where each set corresponds to a vertex, each element corresponds to a hyperedge.
To solve the set packing problem, we just remove the rank 2 restriction of an edge tensor
\begin{equation}
    B_{v,w,\ldots, z} = \begin{cases}
        1, & v+w+\ldots+z\leq 1,\\
        0, & otherwise
    \end{cases}
\end{equation}

@@

We also computed independence polynomial rigorously for these lattices in our Github repo\blue{add link to the repo}.
